FraCaS: Temporal Analysis

### 2020-10-06 paper2020/paper.tex
wibble
kill some more todos and clarify others

The semantics of tense and aspect has been a long standing issue in
the study of formal semantics since the early days of Montague Grammar
and a number of different ideas have been put forth on the table to
deal with them throughout the years. Recent proposals include the
works of the following authors:
\citet{dowty:2012,prior:2003,steedman_productions_2000,higginbotham:2009,fernando:2015}.
%
The
semantics of tense and aspect have been also considered in the study of
Natural Language Inference (NLI). The various datasets
for NLI that have been proposed by the years contain examples that
have some implicit or explicit reliance on inferences related to tense
and aspect. One of the early datasets used to test logical approaches,
the FraCaS test suite~\citep{cooper:1996} contains a whole section
dedicated to temporal and aspectual inference (section 7 of the dataset). This part
of the FraCaS test suite has been difficult to tackle. That is,
so far, no computational system has been capable to deal with it in its entirety: when
authors report accuracy over the FraCaS test suite they
skip this section.
%
In fact, they also often skip the anaphora and ellipsis sections, the
exception being the system presented by
\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019},
which includes support for anaphora and ellipsis but still omit the
temporal section.\footnote{\citet{maccartney:2007} have made a run against the whole test suite, but upon a closer look, this is not really the case, as they do not deal with multi-premise cases. For example, only 36/75 cases in the temporal section are treated. The general accuracy of the system is .59 and .61 for the temporal section. Our system, as shown \cref{tab:results}, presents considerable improvements and coverage over this system.}
%
In this paper, we take up the challenge of providing a
computationally viable account of tense and aspect to deal with the
  section 7 of the FraCaS test suite. Our account is not meant
to be a theoretically extensive account of tense and aspect, but rather an
account that is driven by the need to cover the test suite in a way
that is general enough to capture the test suite examples, \emph{while still covering the rest of the FraCaS test suite}.
%

@@

The paper is structured as follows: in \cref{sec:background}, we give
a brief summary of the computational frameworks whose various
subsystems rely on. In particular, the Grammatical Framework is used
to construct the syntactic parser, the Coq proof assistant checks all
the reasoning and a monad-based dynamic semantics deals with
Montague-style semantics, and references (anaphora).  We also provide
some brief remarks on temporal semantics. In \cref{sec:our-semantics},
we discuss the main aspects of the compositional semantics of our
system, using various examples from the FraCaS suite to illustrate its
effectiveness. In \cref{sec:results}, we evaluate how our system
performs with respect to the FraCaS suite. We ran the system across
the whole suite: our system is thus the first which is capable of
handling the complete FraCaS test suite.  Yet, we are interested in
particular in the performance on the temporal section. In
\cref{sec:conclusions}, we conclude and discuss avenues for future
work.

@@

This paper concludes our work in handling the whole of the FraCaS test
suite using a combination of a number of well-studied tools: type
theory, parsing using the Grammatical Framework (GF), Monadic Dynamic
Semantics and proof assistant technology (Coq). It achieves an
accuracy of 0.71 on the Temporal Section and 0.78 overall.  One of the
things to be looked at is fixing the issues associated with parts of
the test suite that ``broke'' when the temporal analysis was
introduced. Some of these have been already mentioned: interaction of
the temporal variables with group readings and negation.

>>>>

The semantics of tense and aspect has been a long standing issue in
the study of formal semantics since the early days of Montague Grammar
and a number of different ideas have been put forth on the table to
deal with them throughout the years. Recent proposals include the
works of the following authors:
\citet{dowty:2012,prior:2003,steedman_productions_2000,higginbotham:2009,fernando:2015}.
%
The
semantics of tense and aspect have been also considered in the study of
Natural Language Inference (NLI). The various datasets
for NLI that have been proposed by the years contain examples that
have some implicit or explicit reliance on inferences related to tense
and aspect. One of the early datasets used to test logical approaches,
the FraCaS test suite~\citep{cooper:1996} contains a whole section
dedicated to temporal and aspectual inference (section 7 of the dataset). This part
of the FraCaS test suite has been difficult to tackle. That is,
so far, no computational system has been capable to deal with it in its entirety: when
authors report accuracy over the FraCaS test suite they
skip this section.
%
In fact, they also often skip the anaphora and ellipsis sections, the
exception being the system presented by
\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019},
which includes support for anaphora and ellipsis but still omit the
temporal section.\jp{check todo in footnote}\footnote{One can consider that \citet{maccartney:2007} have made a run against the whole test suite. However, they do not deal with multi-premise cases. Consequently only 36/75 cases in the temporal section are treated{TODO: jp: treated or attempted? I suppose attempted (otherwise .61 accuracy makes no sense)}. The general accuracy of the system is .59, and .61 for the temporal section. Our system, as shown \cref{tab:results}, presents considerable improvements in coverage and accuracy over that of \citeauthor{maccartney:2007}.}
%
In this paper, we take up the challenge of providing a
computationally viable account of tense and aspect to deal with the
  section 7 of the FraCaS test suite. Our account is not meant
to be a theoretically extensive account of tense and aspect, but rather an
account that is driven by the need to cover the test suite in a way
that is general enough to capture the test suite examples, \emph{while still covering the rest of the FraCaS test suite}.
%

@@

The paper is structured as follows: in \cref{sec:background}, we give
a brief summary of the computational frameworks whose various
subsystems rely on. In particular, the Grammatical Framework is used
to construct the syntactic parser, the Coq proof assistant checks all
the reasoning and a monad-based dynamic semantics deals with
Montague-style semantics, and references (anaphora).  We also provide
some brief remarks on temporal semantics. In \cref{sec:our-semantics},
we discuss the main aspects of the compositional semantics of our
system, using various examples from the FraCaS suite to illustrate its
effectiveness. In \cref{sec:results}, we evaluate how our system
performs with respect to the FraCaS suite. We ran the system across
the whole suite: our system is thus the first which is capable of
handling the complete FraCaS test suite.  Yet, we are interested in
particular in the performance on the temporal section. In
\cref{sec:conclusion}, we conclude and discuss avenues for future
work.

@@

\cite{bernardy_wide-coverage_2019} presented a
logical system for handling 8 of the 9 sections of the FRACAS test
suite, but excluded section 7, suggesting that it requires many
examples that require ad hoc treatment. Here, we take up this
challenge and show that a system similar to theirs can be extended to
cover the remaining section of the test suite.
%
We use the same combination of a number of well-studied tools: type
theory, parsing using the Grammatical Framework (GF), Monadic Dynamic
Semantics and proof assistant technology (Coq). The system achieves an
accuracy of 0.71 on the Temporal Section and 0.78 overall.  One of the
things to be looked at is fixing the issues associated with parts of
the test suite that ``broke'' when the temporal analysis was
introduced. Some of these have been already mentioned: interaction of
the temporal variables with group readings and negation.
